---
title: "CreateSeuratObject_sybb412project"
output: html_document
author:
  affiliation: 'Case Western Reserve University'
  email: 'jxi98@case.edu'
  name: 'Jeeda Ismail'
date: '`r format(Sys.Date())`'

---

NOTE: this workflow is based on the scripts provided at https://github.com/czbiohub/scell_lung_adenocarcinoma

---


Description
===========

This file shows our workflow to process scRNA-Seq raw data and metadata data in Seurat and then create Seurat Object. We will:  

* Load libraries
* Load in the data
* Create Seurat object
* QC and filtering
* Save filtered Seurat object


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

Load libraries
```{r setup, include=FALSE}
library(useful)
library(Seurat)
library(dplyr)
library(Matrix)
```

Set the full path to our directory
```{r set wd path}
rm(list=ls())
# dir <- "/myVolume/scell_lung_adenocarcinoma/"
dir <- "C:/Users/jeedi/OneDrive - case.edu/Desktop/CWRU/Spring 2021/SYBB 412/Project_SYBB412/"
# dir <- "~/Downloads/scell_lung_adenocarcinoma/"
```

##Loading Data

Load data
```{r load data}

raw.data <- read.csv(paste(dir,"Data_input/S01_datafinal.csv", sep=""), header=T, row.names = 1)
dim(raw.data)
head(colnames(raw.data))
```


Load metadata
```{r load metadata}
metadata <- read.csv(paste(dir,"Data_input/S01_metacells.csv", sep=""), row.names=1, header=T)
head(metadata)
```

Save both objects as RData object 
```{r save object}
save(list=c("metadata","raw.data","dir"), file=paste(dir,"Data_input/S01_Data_and_metadata.RData", sep=""))
```

##Creating Seurat Object

Load RData object 
```{r}
rm(list=ls())
dir <- "C:/Users/jeedi/OneDrive - case.edu/Desktop/CWRU/Spring 2021/SYBB 412/Project_SYBB412/"
load(file = paste(dir,"Data_input/S01_Data_and_metadata.RData", sep=""))
```

Find ERCC's, compute the percent ERCC, and drop them from the raw data.
```{r}
erccs <- grep(pattern = "^ERCC-", x = rownames(x = raw.data), value = TRUE)
percent.ercc <- Matrix::colSums(raw.data[erccs, ])/Matrix::colSums(raw.data)
ercc.index <- grep(pattern = "^ERCC-", x = rownames(x = raw.data), value = FALSE)
raw.data <- raw.data[-ercc.index,]
dim(raw.data)
```

Create the Seurat object with all the data (unfiltered)
```{r}
memory.size(max=60000) #jismail: increasing my RAM memory to run main_tiss

main_tiss <- CreateSeuratObject(counts = raw.data)
# add rownames to metadta 
row.names(metadata) <- metadata$cell_id
# add metadata to Seurat object 
main_tiss <- AddMetaData(object = main_tiss, metadata = metadata)
main_tiss <- AddMetaData(object = main_tiss, percent.ercc, col.name = "percent.ercc")
# Head to check
head(main_tiss@meta.data)
```

Save prefiltered Seurat object 
```{r}
save(main_tiss, file=paste(dir,"Data_input/S02_Main_Seurat_object_prefilter.RData", sep=""))
# load(paste(dir,"Data_input/Main_Seurat_object_prefilter.RData", sep=""))
```

#QC and Filtering


Calculate percent ribosomal genes and add to metadata
```{r}
ribo.genes <- grep(pattern = "^RP[SL][[:digit:]]", x = rownames(x = main_tiss@assays$RNA@data), value = TRUE)
percent.ribo <- Matrix::colSums(main_tiss@assays$RNA@counts[ribo.genes, ])/Matrix::colSums(main_tiss@assays$RNA@data)
main_tiss <- AddMetaData(object = main_tiss, metadata = percent.ribo, col.name = "percent.ribo")
main_tiss
```

Filter cells so that remaining cells have nGenes >= 500 and nReads >= 50000
```{r}
main_tiss_filtered <- subset(x=main_tiss, subset = nCount_RNA > 50000 & nFeature_RNA > 500)
main_tiss_filtered
```

Save filtered Seurat object
```{r}
save(main_tiss_filtered, file=paste(dir,"Data_input/S02_Main_Seurat_object_filtered.RData", sep=""))
```


